Spoken Corpus of Karel Mako\v{n}

### 2018-04-13 2018-clarin-pisa/makon-data.tex
fix hrad nitra's year
add figure for teresa-year correlation
move things around
shortening and polishing clarin2018

\begin{document}
\maketitle
\begin{abstract}
  We present a detailed view on the spoken corpus of Karel Mako\v{n}. This paper
  briefly outlines the system for processing the corpus, which has been
  published in a previous work, and elaborates on the corpus, the process of its
  acquisition and its annotation.
\end{abstract}

@@

The spoken corpus of Karel Mako\v{n} is a collection of talks given in a circle
of friends in the course of late 60's or early 70's till 1991. The recordings
have been kept on magnetophone tapes until their digitization that took place
between 2010 and 2012. A complete transcription was obtained using a dedicated
ASR system and the work of the community around K.M's legacy. The corpus is
about 1000 hours in total length, of which about 66 have been transcribed
manually.

@@

The actual content of the corpus with respect to topics, references and
statements, is on one hand quite well known firstly because there are many books
by the author with descriptive indexes like any other book, which largely
correspond to what is said in the talks, secondly because the the domain stays
more or less consistent across the whole set. On the other hand, it is only
known vaguely and a systematic effort to analyze it is yet to be carried out.

\section{Content of the Corpus}

\subsection{Karel Mako\v{n}}

@@

This culminated in the age of 17, in a perceived life crissis, when he read in a
book that ``this life is the bridge to the eternal life''. The thought
captivated him and gave him a new life direction, which also led to several
ecstasies.

The next culmination happened in 1939, after nine years of constant prayer for a
stronger love to God, when he was doported to the nazi concentration camp in
Sachsenhausen, as a university student. Here, he was promised certain death by a
Gestapo warden. As the nazi was approaching K.M. to kill him, he gave up his
life completely to God and the attacker turned around in awe and fled.

@@

\subsection{Giving and Recording the Talks}

@@

One of the fortunate consequences of the seclusion of the talks is that there is
very little background noise. Also, most of the recordings
% TODO: kolik je od Elgra?
have been taken by a single person who used the available state-of-the-art
amateur recording technology and methods. He was also very systematic in
labeling and archiving the recordings. As a result, the majority of the base for
digitization was a set of reel-to-reel tapes labeled with ordinal indexes and a
set of cassettes labeled by year and ordinal index. The reel-to-reel tapes have
been acquired in a generous speed of 9cm/second, which granted them a very good
shield against the inevitable aging.

\subsection{Topics}

A list of recurring topics and their identification in the sound files is a
point of future work.

We can take the current transcription %and the few notes some listeners made
%about topics and their alignment with the audio
as a source for a baseline.
Topic is inherently a vague concept %, which I shall address when I focus on it
but for the baseline evaluation, we can take a look at some easily defined cases
like named entities. The ease about them is that we can look for their specific
word form and if it is not found, then it means the transcription is not talking
about it.

I have chosen the following example topics:

\begin{enumerate}
\item{Lazarus,}
\item{Mithra, Mithraism,}
\item{Satan,}
\item{St. Teresa,}
\item{fairy-tale.}
\end{enumerate}

The precision of over 80\% seems quite promising and hints that the current
quality of transcription allows for reasonable searching through the corpus.
However, the hits were all clearly audible. I met none in an acoustically defect
(see the next subsection) file, which leads to the hypothesis that those are not
yet searchable to a distantly satisfying level.

The proclamation at the end of the introduction~\ref{intro}, that the
topics of the talks correspond largely to that of the concurrently written
books, deserves a kind of support. Mako\v{n} translated the book \emph{El
Castillo Interior}, English \emph{The Interior Castle} by St. Teresa of Ávila in
1977. For most recordings, the year of their acquisition is known.
Figure~\ref{fig:teresa-year} shows how many hits for the name of the Saint are
present in the transcription by recording year.

\subsection{Acoustic Quality}

Despite of the relatively good recording technique and mostly quiet environment,
the resulting acoustic quality varies wildly. Many recordings are clearly
intelligible, while some suffer from various defects like
\begin{itemize}
\item{background noise in varying intensity and bandwidth,}
\item{altered speed,}
\item{clipping or faitness,}
\item{echo,}
\item{overlap of two recordings.}
\end{itemize}

Sometimes the quality changes abruptly inside a file, when recording was stopped
and resumed with different conditions. This prevents us from classifying
individual recordings by their acoustic quality.

Expectably, the age of a recording correlates with its quality but other factors
play a significant role also. Most are obvious like the relative position
of the speaker to the microphone, the acoustics of the room, quality of the
recording device and also the tape brand. Unfortunately, none of these can be
said with certainty.

@@

I have focused on two major aspects of processing the material: 1. Making it
accessible to the public and 2. acquiring a complete transcription. To a lesser
extent, I have also worked towards 1. overcoming the acoustic difficulties for
listeners, 2. topically indexing some recordings and 3. enablind a full-text
search over the whole corpus.

@@

We only have a single speaker, a restricted domain, relatively consistent
conditions, an extensive collection of the author's original books,
translations, comments and letters to build a well-fitting language model. So
even very little training data should give satisfactory results and the error
rate should drop rapidly as the training data would increase.

Well, the error rate did drop as the training data increased but the accuracy of
the speech recognition was more of a challenge than I had thought. Fot the most
part, the users were actually using the web interface to provide a new
transcription from scratch, ignoring the way too erroneous output of the ASR,
instead of merely correcting single errors, as was meant.

@@

The word error rate on our test set is currently 42\% but it is probably lower
on clean recordings and also most mistakes are in word endings, so the meaning
is comprehensible. I am currently training a DNN acoustic model, so chances are
the ASR score will soon rise noticeably.

@@

\subsubsection{Mitigating Acoustic Defects}

In a new version of the web interface (paper pending), I have implemented a
user-controllable parametric graphical equalizer. This can help significantly in
the case of narrow-band noise, which occurs on some recordings, predominantly in
the bass range.

This was a cheap solution for a low-hanging fruit but a more holistic approach
is very desirable.

>>>>

\begin{document}
\maketitle
\begin{abstract}
  We present a detailed view on the spoken corpus of Karel Mako\v{n}. This paper
  briefly outlines the system for processing the corpus, which has been
  published in a previous work, and elaborates on the corpus. A baseline for
  topical analysis is presented.
\end{abstract}

@@

The spoken corpus of Karel Mako\v{n} is a collection of talks given in a circle
of friends in the course of late 60's or early 70's till 1991. The recordings
have been kept on magnetophone tapes until their digitization that took place
between 2010 and 2012. A complete transcription was obtained using a dedicated
ASR system and the work of the community around K.M.'s legacy. The corpus is
about 1000 hours in total length, of which about 66 have been transcribed
manually.

@@

The actual content of the corpus with respect to topics, references and
statements, is on one hand quite well known
because the the domain stays
more or less consistent across the whole set. On the other hand, it is only
known vaguely and a systematic effort to analyze it is yet to be carried out.

\section{Author}

@@

This culminated in the age of 17, in a perceived life crissis, when he read in a
book that ``this life is the bridge to the eternal life''. The thought
captivated him and gave him a new life direction, which led to several
ecstasies.

The next culmination happened in 1939, after nine years of constant prayer for a
stronger love to God, when he was deported to the nazi concentration camp in
Sachsenhausen, as a university student. Here, he was promised certain death by a
Gestapo warden. As the nazi was approaching K.M. to kill him, he gave up his
life completely to God and the attacker unexplainably turned around in awe and
fled.

@@

\section{Giving and Recording the Talks}

@@

One of the fortunate consequences of the seclusion of the talks is that there is
very little background noise on most of the recordings. Also, most of them
% TODO: kolik je od Elgra?
have been taken by a single person who used the available state-of-the-art
amateur recording technology and methods. He was also very systematic in
labeling and archiving the recordings. As a result, the majority of the base for
digitization was a set of reel-to-reel tapes labeled with ordinal indexes and a
set of cassettes labeled by year and ordinal index. The reel-to-reel tapes have
been acquired in a generous speed of 9cm/second, which granted them a very good
shield against the inevitable aging.

@@

We don't have the resources needed to manually annotate the corpus like
e.g.~\ref{japsponta}, but there are lay
volunteers willing to work on the project for the cause of tending Mako\v{n}'s
legacy.

A complete transcription is essential or at least helpful for any kind of
further processing.  Therefore,
I have focused on two major aspects of processing the material: 1. Making it
accessible to the public and 2. acquiring a complete transcription.

@@

We only have a single speaker, a restricted domain, relatively consistent
conditions, an extensive collection of the author's original books,
translations, comments and letters to build a fitting language model. So
even very little training data should give satisfactory results and the error
rate should drop rapidly as the training data would increase.

Well, the error rate did drop as the training data increased but the accuracy of
the speech recognition was more of a challenge than I had thought. For the most
part, the users were actually using the web interface to provide a new
transcription from scratch, ignoring the way too erroneous output of the ASR,
instead of merely correcting single errors, as was meant.

@@

The word error rate on our test set is currently 42\% but it is probably lower
on clean recordings and also most mistakes are in word endings, so the meaning
is comprehensible. I am currently working on a DNN acoustic model, so chances
are the ASR score will soon rise.

\subsubsection{Acoustic Quality}

Despite of the relatively good recording technique and mostly quiet environment,
the resulting acoustic quality varies wildly. Many recordings are clearly
intelligible, while some suffer from various defects like
1. background noise in varying intensity and bandwidth,
2. altered speed,
3. clipping or faitness,
4. echo,
5. overlap of two recordings.

The varying acoustic quality is the most serious obstacle to the quality of the
automatic transcription. But it is also detrimental to the manual transcription.

To mitigate this, the web interface features since lately a user-controllable
graphical equalizer, which can help especially in the case of narrow-band noise.

@@

Some limited effort was invested into further processing beyond plain
transcription. I have created a simple search engine over the transcription that
yields deep links into relevant passages in the audio. Elasticsearch was used
for this purpose, with Czech algorithmic stemmer.

Another effort was processing manually written indexes that are available for a
fraction of the recordings. These have been partly hand-written, partly typed
and are aligned using the positions of the device counter. Most of these are
scanned and some limited experiments were carried out to apply OCR. The explicit
mapping of the indexes with corresponding recordings and time-positions is yet
to be done.

\section{Topics}

A list of recurring topics and their identification in the sound files is a
point of future work.
We can take the current transcription %and the few notes some listeners made
%about topics and their alignment with the audio
as a source for a baseline.
Topic is inherently a vague concept %, which I shall address when I focus on it
but for the baseline evaluation, we can take a look at some easily defined cases
like named entities. The ease about them is that we can look for their specific
word form and if it is not found, then it means the transcription is not
covering it.

I have chosen the following example topics:
1. \emph{Lazarus},
2. \emph{Mithra, Mithraism},
3. \emph{Satan},
4. \emph{St. Teresa},
5. \emph{fairy-tale}.

The precision of over 80\% seems quite promising and hints that the current
quality of transcription allows for reasonable searching through the corpus.
However, the hits were all clearly audible. I met none in an acoustically defect
file, which leads to the hypothesis that those are not yet searchable to a
distantly satisfying level.

K.M. was writing, translating and commenting books most of his life. We can
assume that the book he was just working on influenced the topic of his talks. 
For on, he translated the book \emph{El Castillo Interior}, English \emph{The
Interior Castle} by St. Teresa of Ávila in 1988. For most recordings, the year
of their acquisition is known.  Figure~\ref{fig:teresa-year} shows how many hits
for the name of the Saint are present in the transcription by recording year.

\begin{figure}[htpb]
\includegraphics[scale=0.6]{rc/teresa-by-year.eps}
\caption{Number of hits for the query \texttt{terez.*} in the transcription by
alleged year of the recording}
\label{fig:teresa-year}
\end{figure}

The peak around the year 1988 supports the speculation that topics in the talks
correlate with those of books written at the same time.
